Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-18 chi2012.tex
foo bar

We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.  In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.  We used a three stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.  Our results suggested that people preferred to rely upon multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.

Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.  Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

\subsection{Better Supporting the Task}

- are people more thorough
- do they evaluate more sources

\subsection{Causing Less Frustration}

- Gave up with aspects of the tasks later / not at all
- With websites, users gave up on websites and moved on early

\subsection{Keeping an Open Mind}

- Did users rule out choices less/later
- Did that mean they chose "better" options

\subsection{What method did people use?}

- Processes
- Organisation of data/screen
- A vs B, how are they different?

% DAN's  qual analysis of processes
In order to complete the tasks, the users followed a number of different processes. The most common process followed
was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
of these next to the set of choices. They then made their choice based on the value of this tally.

Two contrasting approaches that participants followed was to use the data to rule out candidates, or to use the data to
directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
when using the DataPalette interface, significantly more comparison (and visualisation of values) was performed, with fewer
participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with
mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.

Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
often pointing to the screens to show and verify each choice as they explained them.

\subsection{}

- data integration literature demands 'correct' but perfect is the enemy of the good.

@@

% ``perfect'' be the enemy of the good

>>>>

We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.  In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.  We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.  Our results suggested that people preferred to rely upon multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.

Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.  Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping (retail), social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

\subsection{Supporting more sophisticated heterogeneity}

Supporting better structural heterogeneity:
   Combining fields
   Splitting fields
   Rearranging fields (British dates American dates)

Modeling heterogeneity (?)
   Specificity  
     Education+Jobs vs ``Education'' + ``Jobs''

   Unit conversion (kg -> stones)
   Granularity
     Crimes per area

Working with data values

%% \subsection{Better Supporting the Task}

%% - are people more thorough
%% - do they evaluate more sources

%% \subsection{Causing Less Frustration}

%% - Gave up with aspects of the tasks later / not at all
%% - With websites, users gave up on websites and moved on early

%% \subsection{Keeping an Open Mind}

%% - Did users rule out choices less/later
%% - Did that mean they chose "better" options

%% \subsection{What method did people use?}

%% - Processes
%% - Organisation of data/screen
%% - A vs B, how are they different?

% DAN's  qual analysis of processes
%% In order to complete the tasks, the users followed a number of different processes. The most common process followed
%% was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
%% in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
%% Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
%% some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
%% of these next to the set of choices. They then made their choice based on the value of this tally.

%% Two contrasting approaches that participants followed was to use the data to rule out candidates, or to use the data to
%% directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
%% when using the DataPalette interface, significantly more comparison (and visualisation of values) was performed, with fewer
%% participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
%% able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with
%% mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.

%% Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
%% often pointing to the screens to show and verify each choice as they explained them.

@@

% - data integration literature demands 'correct' but perfect is the enemy of the good.
% ``perfect'' be the enemy of the good
